Modelling the high-voltage grid using open data for Europe and beyond

### 2024-11-14 paper/nature-scientific-data-response-letter/response.tex

\documentclass{response}

\title{Modelling the high-voltage grid using open data for Europe and beyond}
\authors{Bobby Xiong, Davide Fioriti, Fabian Neumann, Iegor Riepin, Tom Brown} 
\journal{Scientific Data --- Nature}
\iteration{1}
\date{\today}

\bibliographystyle{elsarticle-num}

\usepackage{libertinust1math}
\renewcommand{\ttdefault}{\sfdefault}
\usepackage[nameinlink,sort&compress,capitalise,noabbrev]{cleveref}
\usepackage{lipsum}

\begin{document}

\maketitle

We thank the reviewers for their constructive comments and acknowledge the time
and effort they have spent in assessing our work. We have revised the paper
based on the feedback from the reviewers and hope that we can adequately address
their primary concerns.

To follow the revisions made, we have highlighted differences compared with the
previous submitted version of the paper in {\color{blue}{blue}} and
{\color{red}{red}} text in an attached file. Figure numbers in this response letter
refer to the numbering in the revised version of the manuscript.

\section*{Editor comments}

\EC Please provide an additional copy of your manuscript file under 20MB as a \textit{Related Manuscript File} in your next submission for internal use.

\AR We will provide an additional copy of the manuscript file under 20 MB, as requested.

\EC A data descriptor provides a description of a static dataset, rather than any novel methods or computational validation schemes. Please edit the text of your abstract to focus more on the dataset that you have generated, rather than a method for constructing the dataset.

\AR Thank you for your feedback and for helping us to improve the clarity and focus of our manuscript. We have revised the abstract to highlight the dataset itself that we have generated, rather than the methodology for its construction.

\EC Because at least some of your data looks to originate from third parties, please check the following and confirm/state what has been done in your \textit{response to reviewers} document:
\begin{enumerate}
    \item All sources are clearly described in either the main text. For general mentions of services, databases, or other platforms, please name the resource in the main text and provide a URL in ()s. Instructions should be sufficient for the exact input data to be retrieved, so please provide specific links, accession numbers, or the exact search term/query when discussing any data from general databases. For specific datasets downloaded from repositories, or other items with formal metadata, please use a full citation in the reference list containing the relevant DOI. Note this should be the dataset, rather than any relevant publication, however these should be cited as well if required, or if the data was newly extracted from a document.
    \item Confirm all the sources are openly available --- i.e. your re-use or re-distribution is compliant with the terms and conditions/licenses for data sharing. If you cannot see a fully open licence at the source please check with the data provider. Please note it is your legal responsibility to ensure all data have been used or re-distributed appropriately and we cannot support publication of descriptions of data obtained or re-shared without this check.
\end{enumerate}

\AR response

\EC There are multiple versions of your dataset at Zenodo. Please cite the specific accession (citation 29) that is relevant for this work.

\AR response

\EC Please provide further description in the Data Record on the contents of this dataset (e.g. variables used, column headings, file names, etc) and any folder structure if relevant. The purpose of the Data Record section is to describe your data contents at a level of detail so that others may re-use them.

\AR response

\EC The description of the data in this article is minimal, making the interpretation of the provided data files hard and even leading to potential discrepancies between the description and the dataset.

\AR response

\newpage
\section*{Reviewer \#1}

\RC Dear Authors, the workflow and datasets described in this manuscript are very valuable to the energy modelling community. It builds and extends already existing models and datasets. I have some notes:

\AR response

\RC Oemof is missing as an open source model in the introduction.

\AR response

\RC The code will be of more value if it is shared as a stand alone code and not as part of the PyPSA-Eur code.

\AR response

\RC Is there a reason behind clustering nodes in a 5000 meters radius? did you check for sensitivity to this value?

\AR response

\RC What happens to this model if overpass turbo is not maintained anymore?

\AR response

\RC Do you also search for the tag \textit{station}, \textit{sub\_station} and the like? some substation in OSM are not well tagged (i.e. do not all have \textit{substation} as tag).

\AR response

\RC Did you compare the relations data and the cable/line data representation for DC lines? The same question for AC lines.

\AR response

\RC There are small typos in the text, which should be corrected. Best regards

\AR response

\newpage
\section*{Reviewer \#2}

\RC The paper presents a description of several csv datafiles that enable construction of the European high-voltage grid for voltages above 200 kV and a method for obtaining these.

\RC The data is derived from open-street maps and presents new annotations compared to the primary sources via a cleaning procedure and attachment of electrical parameters. As such, the data becomes usable for studies such as energy system modelling, or mere electric power system studies.

\RC The aim of the dataset is to have an openly availably and up-to-date set of lines of the electric power system. Therefore, an automated extraction and cleaning method for data obtained from Open Street Maps is proposed by the authors. The quality of the dataset therefore depends on the Open Street Map contributions. Several heuristics are used to clean the data obtained from Open Street Maps, which are methodologically sound. One of the main assumptions is the use of “standard” transmission line types with associated impedances and nominal powers. With the advent of new transmission line types (e.g., high-temperature low-sag conductors), such an approach may not be suitable for future updates and may require updates to the method.

\RC The technical quality of the dataset is supported in two ways: (i) through comparison against official statistics from ENTSO-E on European transmission grid line lengths and against data obtained from a German TSO, (ii) through a comparison of an optimization outcome against a benchmark system.

\RC The first method supports the technical quality of the dataset and proves that the method of retrieving data (using Open Street Maps) can be more reliable than other data sources.

\AR response

\RC The first conclusion of the second method, that \textit{both grids adequately represent reality} cannot be stated as such as (i) a future scenario is being evaluated, and (ii) none of both grids represents reality. The reviewer therefore suggests removing this statement. Furthermore, the reviewer suggests to provide more information of the benchmarking case (e.g., the techno-economic assumptions projected for the year 2030), as these are not described. It is not possible to determine in how much the outcomes of this validation case study depend on these assumptions, e.g., whether other assumptions would lead to larger deviations. Could the authors clarify this?

\AR response

\RC I am a bit surprised by the statement on the data repository itself: 
\begin{quote}
    The authors of this dataset do not claim correctness, completeness, or any other form of guarantee regarding the information presented.
\end{quote}
Although I understand the reasoning behind this statement, this actually undermines the validation of the dataset provided in this article, and conflicts with the statement that \textit{the grids adequately represents reality}.

\AR response

\RC The descriptions of the data files in the proposed article is minimal, mainly in that the authors claim that csv files are provided. The authors could provide a more-in depth description of (i) the meaning of the separate data files on the repository, (ii) the structure of each data file, (iii) the size of the data file.

\AR response

\RC As the description of the data file is minimal, it is not possible to verify whether the data files submitted to the repository are complete. First, a discrepancy between the description of the datafiles (e.g., 400 kV is mapped onto 380 kV) was noted, as 400 kV is still included in the \textit{voltage} column of buses.csv. Could the authors please verify this? Second, when verifying e.g., \textit{lines.csv}, it is not possible to find, e.g., electrical parameters such as nominal current or impedance. Could the authors include an improved description of this data in the article?

\AR response \cite{test}

\newpage
\bibliography{references.bib}

\end{document}

### 2024-11-14 paper/nature-scientific-data-resubmission/main.tex
Added initial paper drafts.

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta}
\usepackage{ulem}  % for striking through text
\usepackage{glossaries}
\usepackage{csvsimple}
\usepackage{bm} % Bold math
\usepackage{todonotes}
\usepackage[version=4]{mhchem}
\usepackage{eurosym}
\usepackage[author={Bobby Xiong}]{pdfcomment}
\usepackage{makecell} % for line breaks in table cells
\usepackage{pdflscape}
\usepackage{array}
\usepackage{hyperref}
\glsdisablehyper % Disable hyperlinks for glossary entries
\let\autocite\cite
% \linenumbers

% Redefine DOI command
\makeatletter
\def\doi#1{%
  \href{https://doi.org/#1}{\textnormal{#1}}%
}
\makeatother

@@

\begin{abstract}
This paper provides the background, methodology and validation for constructing a representation of the European high-voltage grid, including and above 200 kV, based on public data provided by OpenStreetMap. The model-independent grid dataset is published under the Open Data Commons Open Database (ODbL 1.0) licence and can be used for large-scale electricity as well as energy system modelling. The dataset and workflow are provided as part of PyPSA-Eur -- an open-source, sector-coupled optimisation model of the European energy system. By integrating with the codebase for initiatives such as PyPSA-Earth, the value of open and maintainable high-voltage grid data extends to the global context. By accessing the latest data through the the Overpass turbo API, the dataset can be easily reconstructed and updated within minutes. To assess the data quality, this paper further compares the dataset with official statistics and representative model runs using PyPSA-Eur based on different electricity grid representations.
\end{abstract}
\begin{document}

\flushbottom
\maketitle

@@

\section*{Background \& Summary}
\pdfbookmark[1]{Background \& Summary}{background-and-summary}
Energy system models are indispensable tools in today's world in order to understand the complex interactions between energy sources, technologies, policies, and markets. They are used by researchers, industry and policy makers to enable informed decision-making in the transition to a net-zero energy system. However, conclusions drawn from such models are only as good as the underlying data and assumptions. Especially the representation of existing energy infrastructure, such as the electricity grid, can have a deciding impact on future investments derived from such models.\autocite{horschRoleSpatialScale2017} While \glspl{tso} have their own information on the high-voltage grid, this data is often not publicly available to the level of detail needed for academic research purposes. Official institutions like the \gls{entsoe} provide an online map \autocite{entso-eENTSOETransmissionSystem} of the European high-voltage grid. There are however, several limitations typical for these sources: i) there is no underlying, topologically connected dataset, ii) it is not released under an open licence, iii) nor updated frequently, and iv) its geographic detail is limited or highly stylised.

@@

Compared to previous implementations in the global modal PyPSA-Earth,\autocite{parzenPyPSAEarthNewGlobal2023} we significantly improve the work in speed and data quality by taking advantage of the topological, electrical and geographical information available for Europe in \gls{osm}. Given the generic structure of the developed workflow, it can be easily applied to other regions and fed back to the global PyPSA-Earth project. However, the output will directly depend on the \acrshort{osm} data quality for a particular region (e.g. whether data on the substation's geometric footprint is available). To fill in missing data, we introduce cleaning process that yields a representation of the European high-voltage grid. 
We benchmark the processed data against country-level statistics provided by ENTSO-E, concluding that \acrshort{osm} data coverage of the European high-voltage grid is high or even close to complete. These improvements will also contribute to the quality of transmission grids modelled on a global scale in PyPSA-Earth. 
\section*{Methods}
\pdfbookmark[1]{Methods}{methods}
PyPSA-Eur is a spatially and temporally highly resolved, open-source, sector-coupled linear optimisation model that covers the European continent.\autocite{horschPyPSAEurOpenOptimisation2018} The model is build on top of the open-source toolbox PyPSA \autocite{brownPyPSAPythonPower2018} and is suited for operational as well as expansion studies (transmission, generation, and storages). The model includes a stock of existing power plants (processed with the tool powerplanmatching\autocite{gotzensPerformingEnergyModelling2019}) as well as renewable potentials and availability time series (processed with atlite\autocite{hofmannAtliteLightweightPython2021}).
Throughout the last decade, PyPSA-Eur has gained a large user base from academia, industry, and policy makers alike and has been used in a variety of studies.\autocite{neumannPotentialRoleHydrogen2023,victoriaSpeedTechnologicalTransformations2022,brownUltralongdurationEnergyStorage2023,glaumOffshorePowerHydrogen2024,riepinMeansCostsSystemlevel2024,rahdanDistributedPhotovoltaicsProvides2024,grochowiczUsingPowerSystem2024,transnetbwStromnetz2050Studie2022} Other open-source models exist, one notable being OSeMOSYS Global,\autocite{barnesOSeMOSYSGlobalOpensource2022} however it lacks the detailed geographical as well as electrical representation of the transmission grid that PyPSA-Eur provides. With the integration into PyPSA-Eur, we also enable compatibility with additional functions already implemented into the model, such as, but not limited to, the option to enable dynamic line rating\autocite{glaumLeveragingExistingGerman2023} and adding projects under planning (e.g. European Ten-Year Network Development Plan\autocite{entso-eTenYearNetworkDevelopment2020} and the German Network Development Plan\autocite{bnetzaBestaetigungNetzentwicklungsplanStrom2024}).

@@

\subsection*{Step 1 --- Retrieving the OSM data}
\pdfbookmark[2]{Step 1 - Retrieving the OSM data}{methods-step1}
In \gls{osm}, geographical data is stored in `nodes', `ways', and `relations'. Being the simplest data type, nodes are defined by coordinates and associated parameters. Ways are geometric line strings that connect a set of nodes. Relations can contain nodes, ways, a combination of either or other relations. \autocite{medjroubiOpenDataPower2017} For the purpose of this work, we extract ways for obtaining the outline of substations and \acrshort{ac} power lines and cables. To obtain \acrshort{dc} projects (usually cables, hereafter refered to as \acrshort{dc} links), we use relations. There are two main reasons for this differentiation: i) relations are the more complex data type, coverage for \acrshort{ac} lines and cables is still scarce for Europe, whereas for \acrshort{dc} links, given that there are much fewer of them in Europe, coverage is close to complete (see Table \ref{tab:dc_links}); ii) Some \acrshort{dc} links contain multiple components, accessing relations allows us to efficiently aggregate and simplify the links for the purpose of static energy system modelling.
First, we retrieve the raw data using the Overpass turbo \gls{api}\autocite{raiferOverpassTurbo2024} (Figure \ref{fig:all_steps}.a). Specifically, we query the \gls{osm} database for electricity grid related features (see Table \ref{tab:overpass_queries}). Note that Overpass turbo has a limit on the size and number of requests for each query and provides the \gls{api} under a fair use policy. To avoid unnecessary load and re-use of data, we provide a prepared transmission grid for for download via Zenodo. \autocite{xiongPrebuiltElectricityNetwork2024}

\subsection*{Step 2 --- Cleaning the OSM data}
\pdfbookmark[2]{Step 2 - Cleaning the OSM data}{methods-step2}
While \gls{osm} provides a rich dataset, it is not directly usable for energy system modelling. Next to geospatial coordinates, \gls{osm} includes tags that provide feature-specific additional information. Depending on the individual feature, data may however contain noise, be incomplete, or inconsistent. After importing the retrieved raw data into a pandas dataframe,\autocite{mckinneyDataStructuresStatistical2010} we apply a series of steps including heuristics to clean and fill in the missing information (see example in Table \ref{tab:aclines_example1}). We then use the power of geopandas\autocite{jordahlGeopandasGeopandasV02020} to perform geospatial operations (including but not limited to spatial joins, intersections, buffering, etc.).

@@

\subsubsection*{AC lines and cables} In a first step, we clean the tag columns to only contain the correct data type and unit, as shown in Table \ref{tab:aclines_params} for \acrshort{ac} power lines and cables. The minimum parameters that need to be given for a particular line or cable are `voltage' (in \si{\volt}) and `power' (string: `line' or `cable').
% 
\begin{table}[ht]
    \centering
    \begin{tabular}{|l|c|c|}
    \hline
    \textbf{Tag} & \textbf{Data type} & \textbf{Example} \\
    \hline
    cables & numeric & 9 \\
    \hline
    circuits & numeric & 3 \\
    \hline
    frequency & numeric & \SI{50}{\hertz} \\
    \hline
    power & string & line \\
    \hline
    voltage & numeric & \SI{380000}{\volt} \\
    \hline
    \end{tabular}
    \caption{Key tags/parameters for \acrshort{ac} power lines and cables.}
    \label{tab:aclines_params} 
\end{table}

@@

\subsection*{Step 3 --- Building the OSM network}
\pdfbookmark[2]{Step 3 - Building the OSM network}{methods-step3}
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\linewidth]{figures/fig_build_osm_network.pdf}
    \caption{Illustration of building a topologically connected grid.}
    \label{fig:build_osm_network}
\end{figure}

@@

@@

\subsection*{Step 4 --- Creating a PyPSA-ready base network}
\pdfbookmark[2]{Step 4 -Creating a PyPSA-ready base network}{methods-step4}
In a final step, we create a network ready for modelling within PyPSA and PyPSA-Eur. The benefit of our our dataset is that it is provided in .csv format and easily readable. Hence, it may potentially be used outside our given use case, directly or with small adaptations.

@@

\section*{Data Records}
\pdfbookmark[1]{Data Records}{data-records}
The compiled representation of the European high-voltage grid (Figure \ref{fig:osm_map}) is hosted online in .csv format and can be downloaded via the Zenodo repository.\autocite{xiongPrebuiltElectricityNetwork2024} The dataset includes the geographical scope of the ENTSO-E member states (without Cyprus, Iceland, Kosovo, and Turkey). 
We continuously update the dataset as the underlying \gls{osm} input data and the workflow is improved. As of the submission of this paper, the resulting network (Figure \ref{fig:osm_map}) contains 5848 buses, 7320 \acrshort{ac} lines and cables, 36 simplified/aggregated \acrshort{dc} links (and converters at their endpoints, respectively), and 1059 \acrshort{ac} transformers, comprising a total of \SI{261757}{\kilo\meter} in total route length.

@@

\newpage
\newpage
\section*{Technical Validation}
\pdfbookmark[1]{Technical Validation}{technical-validation}
We perform two validation steps for assessing the quality of the dataset. First, we compare the dataset with official inventory statistics provided by \acrshort{entsoe}. Second, we compare the results of an representative PyPSA-Eur model instance based on the two high-voltage grid datasets: OSM (presented in \ref{fig:osm_map}) and an extract from the online ENTSO-E map using GridKit tool (referred to as ‘ENTSO-E map’).\autocite{wiegmansGridKitExtractENTSOE2016} This network is currently being used by numerous PyPSA-Eur users, and is hence a good reference for comparison. In Figure \ref{fig:osm_vs_50Hertz_sgm}, we further provide a comparison between a geo-referenced, official dataset by 50Hertz\autocite{50hertzStaticGridModel2022} and the \acrshort{osm}-based grid of the region, respectively.

\subsection*{Comparison with ENTSO-E statistics and map}
\pdfbookmark[2]{Comparison with ENTSO-E statistics and map}{technical-validation-entsoe}
Based on \acrshort{entsoe}'s 2023 inventory of transmission,\autocite{entso-eInventoryTransmission20232024} we first compare the total route (a) and circuit lengths (b) of \acrshort{ac} lines and cables on a per country level (Figure \ref{fig:bar_lengths}.). Note that the inventory does not include all statistics for each country, i.e. route lengths are missing for Bosnia and Herzegovina, Switzerland, and Great Britain, while circuit lengths are missing for Montenegro and North Macedonia. While Ukraine and the Republic of Moldova have joined \acrshort{entsoe} on 1 January 2024 and 22 November 2023 as full and observing members, respectively, their inventory are not yet included in the dataset. For these two countries, we take reference data from third party sources.\autocite{cigrePowerSystemUkraine2018,globaldataTopFiveTransmission2023,moldelectricaTechnicalEconomicIndicators2023}

@@

@@

\section*{Usage Notes}
\pdfbookmark[1]{Usage Notes}{usage-notes}
The published dataset is provided under the \gls{odbl} 1.0 licence. Geoinformation is encoded in the WGS84 (EPSG:4326) coordinate system. Although the dataset and workflow are provided as part of PyPSA-Eur, they are also suitable for a wide range of other applications, including network analyses, power flow calculations, as well as an input for other energy energy system models and frameworks. Note that further validation and testing is needed for purposes outside the original scope of this work.

>>>>

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta}
\usepackage{graphicx}
\usepackage{ulem}  % for striking through text
\usepackage[table]{xcolor}
\usepackage{glossaries}
\usepackage{csvsimple}
\usepackage{bm} % Bold math
\usepackage{todonotes}
\usepackage[version=4]{mhchem}
\usepackage{eurosym}
\usepackage[author={Bobby Xiong}]{pdfcomment}
\usepackage{makecell} % for line breaks in table cells
\usepackage{pdflscape}
\usepackage{array}
\glsdisablehyper % Disable hyperlinks for glossary entries
\let\autocite\cite
\linenumbers

@@

\begin{abstract}
% \pdfcomment{This is a manuscript template for Data Descriptor submissions to \emph{Scientific Data} (\href{http://www.nature.com/scientificdata}{http://www.nature.com/scientificdata}). The abstract must be no longer than 170 words, and should succinctly describe the study, the assay(s) performed, the resulting data, and the reuse potential, but should not make any claims regarding new scientific findings. No references are allowed in this section.}
This paper provides the background, methodology and validation for constructing a representation of the European high-voltage grid, including and above 200 kV, based on public data provided by OpenStreetMap. The model-independent grid dataset is published under the Open Data Commons Open Database (ODbL 1.0) licence and can be used for large-scale electricity as well as energy system modelling. The dataset and workflow are provided as part of PyPSA-Eur -- an open-source, sector-coupled optimisation model of the European energy system. By integrating with the codebase for initiatives such as PyPSA-Earth, the value of open and maintainable high-voltage grid data extends to the global context. By accessing the latest data through the the Overpass turbo API, the dataset can be easily reconstructed and updated within minutes. To assess the data quality, this paper further compares the dataset with official statistics and representative model runs using PyPSA-Eur based on different electricity grid representations.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
%  Click the title above to edit the author information and abstract

@@

% \pdfcomment{Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists or tables should be included. Structure of the main text is provided below.}
\section*{Background \& Summary}
% \pdfcomment{(700 words maximum) An overview of the study design, the assay(s) performed, and the created data, including any background information needed to put this study in the context of previous work and the literature. The section should also briefly outline the broader goals that motivated the creation of this dataset and the potential reuse value. We also encourage authors to include a figure that provides a schematic overview of the study and assay(s) design. The Background \& Summary should not include subheadings. This section and the other main body sections of the manuscript should include citations to the literature as needed.}
Energy system models are indispensable tools in today's world in order to understand the complex interactions between energy sources, technologies, policies, and markets. They are used by researchers, industry and policy makers to enable informed decision-making in the transition to a net-zero energy system. However, conclusions drawn from such models are only as good as the underlying data and assumptions. Especially the representation of existing energy infrastructure, such as the electricity grid, can have a deciding impact on future investments derived from such models.\autocite{horschRoleSpatialScale2017} While \glspl{tso} have their own information on the high-voltage grid, this data is often not publicly available to the level of detail needed for academic research purposes. Official institutions like the \gls{entsoe} provide an online map \autocite{entso-eENTSOETransmissionSystem} of the European high-voltage grid. There are however, several limitations typical for these sources: i) there is no underlying, topologically connected dataset, ii) it is not released under an open licence, iii) nor updated frequently, and iv) its geographic detail is limited or highly stylised.

@@

Compared to previous implementations in the global modal PyPSA-Earth,\autocite{parzenPyPSAEarthNewGlobal2023} we significantly improve the work in speed and data quality by taking advantage of the topological, electrical and geographical information available for Europe in \gls{osm}. Given the generic structure of the developed workflow, it can be easily applied to other regions and fed back to the global PyPSA-Earth project. However, the output will directly depend on the \acrshort{osm} data quality for a particular region (e.g. whether data on the substation's geometric footprint is available). To fill in missing data, we introduce cleaning process that yields a representation of the European high-voltage grid. 
We benchmark the processed data against country-level statistics provided by ENTSO-E, concluding that \acrshort{osm} data coverage of the European high-voltage grid is high or even close to complete. These improvements will also contribute to the quality of transmission grids modelled on a global scale in PyPSA-Earth. 
\section*{Methods}
% \pdfcomment{The Methods should include detailed text describing any steps or procedures used in producing the data, including full descriptions of the experimental design, data acquisition assays, and any computational processing (e.g. normalization, image feature extraction). See the detailed section in our submission guidelines for advice on writing a transparent and reproducible methods section. Related methods should be grouped under corresponding subheadings where possible, and methods should be described in enough detail to allow other researchers to interpret and repeat, if required, the full study. Specific data outputs should be explicitly referenced via data citation (see Data Records and Citing Data, below).
% Authors should cite previous descriptions of the methods under use, but ideally the method descriptions should be complete enough for others to understand and reproduce the methods and processing steps without referring to associated publications. There is no limit to the length of the Methods section. Subheadings should not be numbered.}
PyPSA-Eur is a spatially and temporally highly resolved, open-source, sector-coupled linear optimisation model that covers the European continent.\autocite{horschPyPSAEurOpenOptimisation2018} The model is build on top of the open-source toolbox PyPSA \autocite{brownPyPSAPythonPower2018} and is suited for operational as well as expansion studies (transmission, generation, and storages). The model includes a stock of existing power plants (processed with the tool powerplanmatching\autocite{gotzensPerformingEnergyModelling2019}) as well as renewable potentials and availability time series (processed with atlite\autocite{hofmannAtliteLightweightPython2021}).
Throughout the last decade, PyPSA-Eur has gained a large user base from academia, industry, and policy makers alike and has been used in a variety of studies.\autocite{neumannPotentialRoleHydrogen2023,victoriaSpeedTechnologicalTransformations2022,brownUltralongdurationEnergyStorage2023,glaumOffshorePowerHydrogen2024,riepinMeansCostsSystemlevel2024,rahdanDistributedPhotovoltaicsProvides2024,grochowiczUsingPowerSystem2024,transnetbwStromnetz2050Studie2022} Other open-source models exist, one notable being OSeMOSYS Global,\autocite{barnesOSeMOSYSGlobalOpensource2022} however it lacks the detailed geographical as well as electrical representation of the transmission grid that PyPSA-Eur provides. With the integration into PyPSA-Eur, we also enable compatibility with additional functions already implemented into the model, such as, but not limited to, the option to enable dynamic line rating\autocite{glaumLeveragingExistingGerman2023} and adding projects under planning (e.g. European Ten-Year Network Development Plan\autocite{entso-eTenYearNetworkDevelopment2020} and the German Network Development Plan\autocite{bnetzaBestaetigungNetzentwicklungsplanStrom2024}).

@@

\subsection*{Step 1 --- Retrieving the \gls{osm} data}
In \gls{osm}, geographical data is stored in `nodes', `ways', and `relations'. Being the simplest data type, nodes are defined by coordinates and associated parameters. Ways are geometric line strings that connect a set of nodes. Relations can contain nodes, ways, a combination of either or other relations. \autocite{medjroubiOpenDataPower2017} For the purpose of this work, we extract ways for obtaining the outline of substations and \acrshort{ac} power lines and cables. To obtain \acrshort{dc} projects (usually cables, hereafter refered to as \acrshort{dc} links), we use relations. There are two main reasons for this differentiation: i) relations are the more complex data type, coverage for \acrshort{ac} lines and cables is still scarce for Europe, whereas for \acrshort{dc} links, given that there are much fewer of them in Europe, coverage is close to complete (see Table \ref{tab:dc_links}); ii) Some \acrshort{dc} links contain multiple components, accessing relations allows us to efficiently aggregate and simplify the links for the purpose of static energy system modelling.
First, we retrieve the raw data using the Overpass turbo \gls{api}\autocite{raiferOverpassTurbo2024} (Figure \ref{fig:all_steps}.a). Specifically, we query the \gls{osm} database for electricity grid related features (see Table \ref{tab:overpass_queries}). Note that Overpass turbo has a limit on the size and number of requests for each query and provides the \gls{api} under a fair use policy. To avoid unnecessary load and re-use of data, we provide a prepared transmission grid for for download via Zenodo. \autocite{xiongPrebuiltElectricityNetwork2024}

\subsection*{Step 2 --- Cleaning the \gls{osm} data}
While \gls{osm} provides a rich dataset, it is not directly usable for energy system modelling. Next to geospatial coordinates, \gls{osm} includes tags that provide feature-specific additional information. Depending on the individual feature, data may however contain noise, be incomplete, or inconsistent. After importing the retrieved raw data into a pandas dataframe,\autocite{mckinneyDataStructuresStatistical2010} we apply a series of steps including heuristics to clean and fill in the missing information (see example in Table \ref{tab:aclines_example1}). We then use the power of geopandas\autocite{jordahlGeopandasGeopandasV02020} to perform geospatial operations (including but not limited to spatial joins, intersections, buffering, etc.).

@@

\subsubsection*{\acrshort{ac} lines and cables} In a first step, we clean the tag columns to only contain the correct data type and unit, as shown in Table \ref{tab:aclines_params} for \acrshort{ac} power lines and cables. The minimum parameters that need to be given for a particular line or cable are `voltage' (in \si{\volt}) and `power' (string: `line' or `cable').
% 
\begin{table}[ht]
    \centering
    \begin{tabular}{|l|c|c|}
    \hline
    \textbf{Tag} & \textbf{Data type} & \textbf{Example} \\
    \hline
    cables & numeric & 9 \\
    \hline
    circuits & numeric & 3 \\
    \hline
    frequency & numeric & \SI{50}{\hertz} \\
    \hline
    power & string & line \\
    \hline
    voltage & numeric & \SI{380000}{\volt} \\
    \hline
    \end{tabular}
    \caption{Key tags/parameters for \acrshort{ac} power lines and cables.}
    \label{tab:aclines_params} 
\end{table}

@@

\subsection*{Step 3 --- Building the \gls{osm} network}
\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\linewidth]{figures/fig_build_osm_network.pdf}
    \caption{Illustration of building a topologically connected grid.}
    \label{fig:build_osm_network}
\end{figure}

@@

% \begin{itemize}
%     \item First, we enhance the original substations dataset by including the endpoints of both \acrshort{ac} and \acrshort{dc} lines. This step ensures consistency between the substation and line datasets by guaranteeing that each line has a corresponding substation at its endpoint. However, as shown in Fig. \ref{fig:step_3}, this process can introduce additional duplicate substations beyond those already present, which need to be addressed.
%     \item Second, as \gls{osm} data are not model-ready, we need to ensure the connectivity of long lines that may pass through multiple substations. When a line path overpasses a substation, the line is divided into two segments thus ensuring the so-called overpassing node is connected to the whole system. In Fig. \ref{fig:step_3}, `line B' is divided into segments `a' and `b' to ensure the connection of the overpassing bus to the whole system.
%     \item Third, we cluster the buses within a radius of \SI{5000}{\meter} to avoid duplicated substations and improve the computational efficiency of the model. This is done by an iterative procedure that assign a bus ID to each susbtation and its neighbours iteratively. If a neighbour within the radius is already assigned an ID, then its value is set to the current bus; otherwise a new ID is generated. This step is repeated for \acrshort{ac} and \acrshort{dc} substations, hence ensuring different IDs for each type. By doing so, the same ID may be assigned to substations with multiple voltage values.
%     \item Forth, resulting substations are clustered by ID, voltage level and AC/DC status and for each cluster an electrical bus is created. Fig. \ref{fig:step_3} shows an example of the procedure.
%     \item Fifth, to properly represent the network, we then add transformers to connect \acrshort{ac} buses with the same ID but different voltage values. On the other hand, for \acrshort{dc} buses we add a AC/DC converter to ensure connection with the closest high-voltage \acrshort{ac} bus.
%     \item Finally, appropriate data formatting is applied to adapt the naming conventions to the PyPSA standard, hence ensuring the compatibility of the network with the PyPSA model, as described in the following stage.
% \end{itemize}

@@

\subsection*{Step 4 --- Creating a PyPSA-ready base network}
In a final step, we create a network ready for modelling within PyPSA and PyPSA-Eur. The benefit of our our dataset is that it is provided in .csv format and easily readable. Hence, it may potentially be used outside our given use case, directly or with small adaptations.

@@

% \subsubsection*{Third-level section}
% Topical subheadings are allowed.

\section*{Data Records}
% \pdfcomment{The Data Records section should be used to explain each data record associated with this work, including the repository where this information is stored, and to provide an overview of the data files and their formats. Each external data record should be cited numerically in the text of this section, for example \autocite{brownSynergiesSectorCoupling2018}, and included in the main reference list as described below. A data citation should also be placed in the subsection of the Methods containing the data-collection or analytical procedure(s) used to derive the corresponding record. Providing a direct link to the dataset may also be helpful to readers (\hyperlink{https://doi.org/10.6084/m9.figshare.853801}{https://doi.org/10.6084/m9.figshare.853801}).
% Tables should be used to support the data records, and should clearly indicate the samples and subjects (study inputs), their provenance, and the experimental manipulations performed on each (please see 'Tables' below). They should also specify the data output resulting from each data-collection or analytical step, should these form part of the archived record.}
The compiled representation of the European high-voltage grid (Figure \ref{fig:osm_map}) is hosted online in .csv format and can be downloaded via the Zenodo repository.\autocite{xiongPrebuiltElectricityNetwork2024} The dataset includes the geographical scope of the ENTSO-E member states (without Cyprus, Iceland, Kosovo, and Turkey). 
We continuously update the dataset as the underlying \gls{osm} input data and the workflow is improved. As of the submission of this paper, the resulting network (Figure \ref{fig:osm_map}) contains 5848 buses, 7320 \acrshort{ac} lines and cables, 36 simplified/aggregated \acrshort{dc} links (and converters at their endpoints, respectively), and 1059 \acrshort{ac} transformers, comprising a total of \SI{261757}{\kilo\meter} in total route length.

@@

\newpage
\newpage
\section*{Technical Validation}
% \pdfcomment{This section presents any experiments or analyses that are needed to support the technical quality of the dataset. This section may be supported by figures and tables, as needed. This is a required section; authors must present information justifying the reliability of their data.}
We perform two validation steps for assessing the quality of the dataset. First, we compare the dataset with official inventory statistics provided by \acrshort{entsoe}. Second, we compare the results of an representative PyPSA-Eur model instance based on the two high-voltage grid datasets: OSM (presented in \ref{fig:osm_map}) and an extract from the online ENTSO-E map using GridKit tool (referred to as ‘ENTSO-E map’).\autocite{wiegmansGridKitExtractENTSOE2016} This network is currently being used by numerous PyPSA-Eur users, and is hence a good reference for comparison. In Figure \ref{fig:osm_vs_50Hertz_sgm}, we further provide a comparison between a geo-referenced, official dataset by 50Hertz\autocite{50hertzStaticGridModel2022} and the \acrshort{osm}-based grid of the region, respectively.

\subsection*{Comparison with \acrshort{entsoe} statistics and map}
Based on \acrshort{entsoe}'s 2023 inventory of transmission,\autocite{entso-eInventoryTransmission20232024} we first compare the total route (a) and circuit lengths (b) of \acrshort{ac} lines and cables on a per country level (Figure \ref{fig:bar_lengths}.). Note that the inventory does not include all statistics for each country, i.e. route lengths are missing for Bosnia and Herzegovina, Switzerland, and Great Britain, while circuit lengths are missing for Montenegro and North Macedonia. While Ukraine and the Republic of Moldova have joined \acrshort{entsoe} on 1 January 2024 and 22 November 2023 as full and observing members, respectively, their inventory are not yet included in the dataset. For these two countries, we take reference data from third party sources.\autocite{cigrePowerSystemUkraine2018,globaldataTopFiveTransmission2023,moldelectricaTechnicalEconomicIndicators2023}

@@

% begin math environment

@@

\section*{Usage Notes}
% The Usage Notes should contain brief instructions to assist other researchers with reuse of the data. This may include discussion of software packages that are suitable for analysing the assay data files, suggested downstream processing steps (e.g. normalization, etc.), or tips for integrating or comparing the data records with other datasets. Authors are encouraged to provide code, programs or data-processing workflows if they may help others understand or use the data. Please see our code availability policy for advice on supplying custom code alongside Data Descriptor manuscripts.

% For studies involving privacy or safety controls on public access to the data, this section should describe in detail these controls, including how authors can apply to access the data, what criteria will be used to determine who may access the data, and any limitations on data use.
